Compositional Certified Resource Bounds

### 2015-04-08 paper/main.tex
compactify
rework the t62 explanation

\begin{figure*}
  \centering
\ifshort{\vspace{-.3cm}
\hspace{-0.4cm}}{}
  \begin{minipage}[b]{7cm}
    \begin{center}
   \begin{lstlisting}[]
void c_down (int x,int y) {
  if (x>y) {tick(1); c_up(x-1,y);}
}
void c_up (int x, int y) {
  if (y+1<x) {tick(1); c_down(x,y+2);}
}
   \end{lstlisting}
\vspace{-2.7ex}
$0.33 + 0.67 |[y,x]|\;\;$ (\code{c\_down(x,y)})\\
$\;\;\;\;\;\;\,0.67 |[y,x]|\;\;$ (\code{c\_up(x,y)})
\\[.4\baselineskip]
      {\bf t39}
    \end{center}
  \end{minipage}
%
\hfill
%
  \begin{minipage}[b]{5cm}
    \begin{center}
   \begin{lstlisting}
for (; l>=8; l-=8)
  /* process one block */
  tick(N);
for (; l>0; l--)
  /* save leftovers */
  tick(1);
   \end{lstlisting}
\vspace{-2.8ex}
$\;\;\;\;\;\;\;\;\;\;\;\;\;\frac{N}{8} |[0,l]|\;\;$ if $N \geq 8$\\
$7\frac{8-N}{8} + \frac{N}{8}|[0,l]|\;\;$ if $N < 8$
\\[.4\baselineskip]
      {\bf t61}
    \end{center}
  \end{minipage}
%
\hfill
%
  \begin{minipage}[b]{5.5cm}
    \begin{center}
\begin{lstlisting}
for (;;) {
  do { l++; tick(1); }
    while (l<h && *);
  do { h--; tick(1); }
    while (h>l && *);
  if (h<=l) break;
  tick(1); /* swap elems. */ }
\end{lstlisting}
$2+3|[l,h]|$
\\[.4\baselineskip]
      {\bf t62}
    \end{center}
  \end{minipage}
\vspace{1ex}
\caption{Example \emph{t39} shows two mutually-recursive functions
  with the computed tick bounds.  Example \emph{t61} and \emph{t62}
  demonstrate the unique compositionality of our system. In \emph{t61},
  $N\geq 0$ is a fixed but arbitrary constant.  The derived constant factors
  are tight.}
  \label{fig:ex_list_2}
\end{figure*}

@@

Example \emph{t62} (\pref{fig:ex_list_2})
is the inner loop of a quick sort implementation in cBench.  More
precisely, it is the partitioning part of the algorithm.  This partition
loop has linear complexity, and feeding into our analysis gives the
correct and tight worst-case bound $2+3|[l,h]|.$  With the potential
method in mind, it is clear that the three loops get their
potential from the same interval $[l,h]$.  However, KoAT fails to
find a bound and LOOPUS derives the quadratic bound
$(h-l-1)^2$.  Following the classical technique, these tools try to find one
linear ranking-function for each loop and combine them multiplicatively
or additively.

@@

In general, the weakening rule can be applied after every syntax directed
rule.  However, it can be left out in practice at some places to
increase the efficiency of the tool.
The weakening operation $\succeq_{\Gamma}$ is defined by the rule
\textsc{Relax}. It is parameterized by a logical context that is
used to gather information on interval sizes.  For example,
{%\small
  \begin{align*}
  P^\text{de} \succeq_{(\cdot)} P^\text{we} &\equiv
 p^\text{we}_{0,10} \leq p^\text{de}_{0,10} + u_{0,10} - v_{0.10} \\
&\land  p^\text{we}_{0} \leq p^\text{de}_{0} - 10 {\cdot} u_{0,10} + 10 {\cdot} v_{0.10} \\
&\land  \forall (\alpha, \beta) \neq (0,10)  . \, p^\text{we}_{\alpha,\beta} \leq p^\text{de}_{\alpha,\beta} \;.
  \end{align*}
}

\noindent
The other rules are syntax directed and applied inductively.  For example,
the outermost expression is a loop, so we use the rule \text{Q:Loop}
at the root of the derivation tree.  At this point, we do not know yet
whether a loop invariant exists.  But we produce the constraints
$Q^\text{lo} = P^\text{lo}$\iffull{ which is short for the following
constraint set.
$$
\begin{array}{lll}
q^\text{lo}_0 = p^\text{lo}_0 &
q^\text{lo}_{x,0}= p^\text{lo}_{x,0} &
q^\text{lo}_{0,x}= p^\text{lo}_{0,x} \\
q^\text{lo}_{x,10}= p^\text{lo}_{x,10} &
q^\text{lo}_{10,x}= p^\text{lo}_{10,x} &
q^\text{lo}_{0,10}= p^\text{lo}_{0,10}
\end{array}
$$}{.}
%
These constraints express the fact that the potential functions before and
after the loop body are equal and thus constitute an invariant.

>>>>

\begin{figure*}
  \centering
\ifshort{\vspace{-.3cm}
\hspace{-0.4cm}}{}
  \begin{minipage}[b]{7cm}
    \begin{center}
   \begin{lstlisting}[]
void c_down (int x,int y) {
  if (x>y) {tick(1); c_up(x-1,y);}
}
void c_up (int x, int y) {
  if (y+1<x) {tick(1); c_down(x,y+2);}
}
   \end{lstlisting}
\vspace{-2.7ex}
$0.33 + 0.67 |[y,x]|\;\;$ (\code{c\_down(x,y)})\\
$\;\;\;\;\;\;\,0.67 |[y,x]|\;\;$ (\code{c\_up(x,y)})
\\[.4\baselineskip]
      {\bf t39}
    \end{center}
  \end{minipage}
%
\hfill
%
  \begin{minipage}[b]{5cm}
    \begin{center}
   \begin{lstlisting}
for (; l>=8; l-=8)
  /* process one block */
  tick(N);
for (; l>0; l--)
  /* save leftovers */
  tick(1);
   \end{lstlisting}
\vspace{-2.8ex}
$\;\;\;\;\;\;\;\;\;\;\;\;\;\frac{N}{8} |[0,l]|\;\;$ if $N \geq 8$\\
$7\frac{8-N}{8} + \frac{N}{8}|[0,l]|\;\;$ if $N < 8$
\\[.4\baselineskip]
      {\bf t61}
    \end{center}
  \end{minipage}
%
\hfill
%
  \begin{minipage}[b]{5.5cm}
    \begin{center}
\begin{lstlisting}
for (;;) {
  do { l++; tick(1); }
    while (l<h && *);
  do { h--; tick(1); }
    while (h>l && *);
  if (h<=l) break;
  tick(1); /* swap elems. */ }
\end{lstlisting}
$2+3|[l,h]|$
\\[.4\baselineskip]
      {\bf t62}
    \end{center}
  \end{minipage}
\vspace{1ex}
\caption{Example \emph{t39} shows two mutually-recursive functions
  with the computed tick bounds.  Example \emph{t61} and \emph{t62}
  demonstrate the unique compositionality of our system. In \emph{t61},
  $N\geq 0$ is a fixed but arbitrary constant. % The derived constant factors
  %are tight.
  }
  \label{fig:ex_list_2}
\end{figure*}

@@

Example \emph{t62} (\pref{fig:ex_list_2})
is the inner loop of a quick sort implementation in cBench.  More
precisely, it is the partitioning part of the algorithm.  This partition
loop has linear complexity, and feeding it to our analysis gives a
worst-case bound $2+3|[l,h]|$.  The bound is not optimal but it can be
refined by rewriting the program. % to let our logical inference
%discover more invariants.
%With the potential
%method in mind, it is clear that the three loops get their
%potential from the same interval $[l,h]$ regardless of their nesting.
With the potential method in mind, we can reason as follows.
%
If $h \geq l$ initially, the cost of the loop is 2.
Otherwise, the cost of each round (at
most 3) is payed using the potential of $[l,h]$ by the first
increment to $l$ because we know that $l<h$.  The two inner loops
can also use $[l,h]$ to pay for their inner costs.
%
KoAT fails to
find a bound and LOOPUS derives the quadratic bound
$(h-l-1)^2$.  Following the classical technique, these tools try to find one
ranking function for each loop and combine them multiplicatively
or additively.

@@

In general, the weakening rule can be applied after every syntax directed
rule.  However, it can be left out in practice at some places to
increase the efficiency of the tool.
The weakening operation $\succeq_{\Gamma}$ is defined by the rule
\textsc{Relax}. It is parameterized by a logical context that is
used to gather information on interval sizes.  For example,
{%\small
  \begin{align*}
  P^\text{de} \succeq_{(\cdot)} P^\text{we} &\equiv
 p^\text{we}_{0,10} \leq p^\text{de}_{0,10} + u_{0,10} - v_{0.10} \\
&\land  p^\text{we}_{0} \leq p^\text{de}_{0} - 10 {\cdot} u_{0,10} + 10 {\cdot} v_{0.10} \\
&\land  \forall (\alpha, \beta) \neq (0,10)  . \, p^\text{we}_{\alpha,\beta} \leq p^\text{de}_{\alpha,\beta} \;.
  \end{align*}
}
The other rules are syntax directed and applied inductively.  For example,
the outermost expression is a loop, so we use the rule \text{Q:Loop}
at the root of the derivation tree.  At this point, we do not know yet
whether a loop invariant exists.  But we produce the constraints
$Q^\text{lo} = P^\text{lo}$\iffull{ which is short for the following
constraint set.
$$
\begin{array}{lll}
q^\text{lo}_0 = p^\text{lo}_0 &
q^\text{lo}_{x,0}= p^\text{lo}_{x,0} &
q^\text{lo}_{0,x}= p^\text{lo}_{0,x} \\
q^\text{lo}_{x,10}= p^\text{lo}_{x,10} &
q^\text{lo}_{10,x}= p^\text{lo}_{10,x} &
q^\text{lo}_{0,10}= p^\text{lo}_{0,10}
\end{array}
$$}{.}
%
These constraints express the fact that the potential functions before and
after the loop body are equal and thus constitute an invariant.
